\LARGE \bf Grupo Bimbo Inventory Demand Data Science Report

### 2024-07-15 src/data-science-project/data_science_project/docs/paper/paper.tex
Paper 100 percent

\subsection{Data Preprocessing}
The preprocessing steps included:
\begin{itemize}
        \item \textbf{Data Cleaning}: Handling missing values and removing duplicates.
        \item \textbf{Feature Engineering}: Creating new features such as lagged variables to capture temporal dependencies and integrating external economic indicators.
        \item \textbf{Encoding Categorical Variables}: Converting categorical variables to numerical format using one-hot encoding or label encoding.
        \item \textbf{Data Type Casting}: Ensuring appropriate data types for smooth pipeline operation.
\end{itemize}

\subsection{Modeling}
Two machine learning models were employed:
\begin{itemize}
        \item XGBoost: A gradient boosting framework optimized for performance.
        \begin{itemize}
                \item $\textbf{Parameters}$: $n\_estimators=200$, $max\_depth=8$, $learning\_rate=0.2$, $subsample=0.9$, $colsample\_bytree=0.8$, $n\_jobs=-1$, and $random\_state=42$.
        \end{itemize}
        \item Random Forest: An ensemble learning method using multiple decision trees.
        \begin{itemize}
                \item $\textbf{Parameters}$: $n\_estimators=110$, $min\_samples\_split=10$, $min\_samples\_leaf=5$, $n\_jobs=-1$, and $random\_state=42$.
        \end{itemize}
\end{itemize}

\subsection{Data Preprocessing}
The preprocessing steps included:
\begin{itemize}
        \item \textbf{Data Cleaning}: Handling missing values and removing duplicates.
        \item \textbf{Feature Engineering}: Creating new features such as lagged variables to capture temporal dependencies and integrating external economic indicators.
        \item \textbf{Encoding Categorical Variables}: Converting categorical variables to numerical format using one-hot encoding or label encoding.
        \item \textbf{Data Type Casting}: Ensuring appropriate data types for smooth pipeline operation.
\end{itemize}

\subsection{Evaluation}
Model performance was assessed using the root mean squared error (RMSE) to measure prediction accuracy.

@@

\begin{figure}[H] 
        \begin{center}
        \centering
        \includegraphics[width=0.5\textwidth]{images/comparison.jpg}
        \caption{Actual vs Predict Adjusted Demand}
        \end{center}
\end{figure}

\begin{figure}[H] 
        \begin{center}
        \centering
        \includegraphics[width=0.5\textwidth]{images/comparisonxg.jpg}
        \caption{Actual vs Predict Adjusted Demand}
        \end{center}
\end{figure}

>>>>

\subsection{Data Preprocessing}
The preprocessing steps undertaken in this study involved several key processes to ensure the data was clean, consistent, and suitable for model training. Below are the detailed steps followed:
\begin{itemize}                
        \item \textbf{Data Cleaning}: 
        \begin{itemize}
                \item Handling Missing Values: Missing values in the dataset were identified and appropriately handled. For numerical columns, missing values were filled with the median of the column, while for categorical columns, the mode was used.        
        \end{itemize}
        \item Outliers Removal: Outliers were detected using z-score and IQR methods and were removed or capped to prevent them from skewing the model results.

        \item Data Transformation:
        \begin{itemize}
                \item Normalization and Scaling: Numerical features were normalized and scaled to ensure all features contribute equally to the model. StandardScaler from sklearn was used to transform features to have zero mean and unit variance.
                \item Encoding Categorical Variables: Categorical variables were converted into numerical values using one-hot encoding for nominal variables and label encoding for ordinal variables.
        \end{itemize}

        \item \textbf{Feature Engineering}: Creating new features such as lagged variables to capture temporal dependencies and integrating external economic indicators.
        \begin{itemize}
                \item Temporal Features: New temporal features such as the day of the week, month, and lag features for past sales data were created to capture temporal dependencies and seasonality in the data.
                \item Interaction Features: Interaction features between different variables were created to capture complex relationships within the data.
                \item External Economic Indicators: The bi-weekly inflation rate and consumer confidence index were merged with the primary dataset to capture the economic conditions impacting product demand.
        \end{itemize}
\end{itemize}

\begin{figure}[H] 
        \begin{center}
        \centering
        \includegraphics[width=0.5\textwidth]{images/output.png}
        \caption{Heatmap Feature Information}
        \end{center}
\end{figure}

\subsection{Modeling}
Two machine learning models were chosen for this study based on their ability to handle large datasets and complex relationships: XGBoost and Random Forest.
\begin{itemize}
        \item XGBoost: XGBoost (Extreme Gradient Boosting) is an efficient and scalable implementation of gradient boosting framework by Friedman et al. It provides parallel tree boosting that solves many data science problems in a fast and accurate way.
        \begin{itemize}
                \item $\textbf{Parameters}$: $n\_estimators=200$, $max\_depth=8$, $learning\_rate=0.2$, $subsample=0.9$, $colsample\_bytree=0.8$, $n\_jobs=-1$, and $random\_state=42$.
        \end{itemize}
        \item Random Forest is an ensemble learning method for classification, regression, and other tasks that operates by constructing a multitude of decision trees at training time.
        \begin{itemize}
                \item $\textbf{Parameters}$: $n\_estimators=110$, $min\_samples\_split=10$, $min\_samples\_leaf=5$, $n\_jobs=-1$, and $random\_state=42$.
        \end{itemize}
\end{itemize}

\subsection{Evaluation}
Model performance was evaluated using the root mean squared error (RMSE) as the primary metric. RMSE is a standard way to measure the error of a model in predicting quantitative data. Additionally, other metrics such as R-squared, Mean Absolute Error (MAE), and Mean Squared Error (MSE) were also considered for a comprehensive evaluation.
\begin{itemize}
        \item XGBoost Performance:

        \begin{itemize}
                \item RMSE: The RMSE for the XGBoost model was X.
                \item Other Metrics:
                \begin{itemize}
                        \item R-squared: The model's R-squared value was Y.
                        \item MAE: The Mean Absolute Error for the model was Z.
                        \item MSE: The Mean Squared Error for the model was W.
                \end{itemize}
        \end{itemize}

        \item Random Forest Performance: is an ensemble learning method for classification, regression, and other tasks that operates by constructing a multitude of decision trees at training time.

        \begin{itemize}
                \item RMSE: The RMSE for the Random Forest model was A.
                \item Other Metrics:
                \begin{itemize}
                        \item R-squared: The model's R-squared value was B.
                        \item MAE: The Mean Absolute Error for the model was C.
                        \item MSE: The Mean Squared Error for the model was D.
                \end{itemize}
        \end{itemize}
\end{itemize}

@@

\begin{figure}[H] 
        \begin{center}
        \centering
        \includegraphics[width=0.5\textwidth]{images/comparison.jpg}
        \caption{Actual vs Predict Adjusted Demand RandomForest}
        \end{center}
\end{figure}

\begin{figure}[H] 
        \begin{center}
        \centering
        \includegraphics[width=0.5\textwidth]{images/comparisonxg.jpg}
        \caption{Actual vs Predict Adjusted Demand XGBoost}
        \end{center}
\end{figure}
